Parallel Computations for Solving Multicriteria Mixed-Integer Optimization Problems

### 2021-03-26 2021/PСT_GergelKozinov/2021_PCT.tex
PCT финальная версия

\documentclass{svproc}

@@

\begin{abstract}
The paper discusses a new approach to solving computationally time-consuming multicriteria optimization problems, in which some variable parameters can only take on discrete values. Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of optimization problems where only continuous parameters are used. All problems of the family are solved simultaneously in the time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations. The algorithms proposed enable parallel computing on high-performance computing systems. Computational experiments confirm that the proposed approach makes it possible to significantly reduce computation volume and time for solving complex multicriteria mixed-integer optimization problems.
\keywords{Multicriteria optimization $\cdot$ mixed-integer optimization problems $\cdot$ methods of criteria scalarization $\cdot$ global optimization $\cdot$ search information $\cdot$ parallel computing $\cdot$ computational experiment}
\end{abstract}

@@

Multicriteria optimization (MCO) problems arise almost every time when the optimal decision need to be chosen when developing complex technical devices and systems. The scale of demand for MCO problems determines the high intensity of research in this area -- see, for example, monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}.

Usually, the solution of a MCO problem is a set of efficient (non-dominated) decisions for which an improvement in values with respect to some criteria cannot be achieved without deteriorating the efficiency with respect to some other criteria. The definition of the whole set of efficient decisions (the Pareto set), on the one hand, may require a large amount of computation and, on the other hand, may be redundant as the analysis of a large number of efficient decisions may require a significant effort for the person who makes the decision (decision maker, DM). Thus, it may be practically justified to find only a relatively small set of efficient decisions that can be formed according to the optimality requirements defined by the DM.

By limiting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the amount of calculations required. However, efficiency criteria may be complex, \textit{multiextremal}, and calculating the values of these criteria can prove \textit{computationally demanding}. In addition, some variables may only take on discrete values. In such cases, MCO problems involve \textbf{significant computational complexity}, which can only be overcome by using high-performance supercomputer systems.

Many different approaches have been proposed for solving MCO problems (see, for example, \cite{c3,c6,c10,c11}.  Most commonly, various methods based on reduction of the vector criterion to a particular scalar function are used \cite{c2,c12}. The number of works on multicriteria mixed-integer problems is however more limited -- in most cases, the issues of discrete parameter analysis are considered in relation to scalar optimization problems (see, for example, reviews \cite{c13,c14}). The widely used deterministic methods for solving problems of this class are usually based on the Branch-and-Bound \cite{c15} or on the Branch-and-Reduce approaches \cite{c16}. A number of meta-heuristic and genetic algorithms are also known and are based, in one way or another, on the random search concept \cite{c17,c18}.

This paper presents the results of research that was carried out to develop highly effective parallel methods of multicriteria optimization making use of all the search information obtained in the course of calculations \cite{c19,c20,c21}. A new contribution to this research area consists in the development of an approach for solving MCO problems, in which some of the varied parameters can only take discrete values. The proposed approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used. All problems of the family are solved simultaneously in the time-shared mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively taking into account the search information obtained in the course of the calculations. The algorithms developed enable efficient parallel computing on high-performance computing systems.

The rest of the paper is organized as follows. In Section \ref{sec:2}, we give the statement of multicriteria optimization problems, present a minimax scheme for scalarization of the vector efficiency criterion and introduce the concept of multistage solution of multicriteria optimization problems. In Section \ref{sec:3}, we present a description of the proposed approach based on the reduction of mixed-integer optimization problems to solving a family of optimization problems with the use of only continuous parameters. This section also describes a dimensionality reduction scheme, which allows multidimensional optimization problems to be reduced to a one-dimensional global search problem. Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems, which is used in the framework of the proposed approach. Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is promising. In conclusion, the results obtained are discussed and possible main directions for continuing the research are outlined.

@@

The problem of multicriteria mixed-integer optimization (MCOmix) can be formulated as follows
\begin{equation}\label{eq:1}
f(y,u)\to \min, y\in D,u \in U,
\end{equation}
where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency in which the varied parameters belong to two different types:
\begin{itemize}
	\item continuous parameters $y=(y_1,y_2,\dots,y_n)$, whose domain of possible values is represented by an $N$-dimensional hyperparallelepiped:
\begin{equation}\label{eq:2}
D=\{ y\in R^n: a_i \leq y_i \leq b_i, 1 \leq i \leq n \}, 
\end{equation}
for specified vectors $a$ and $b$.
	\item discrete parameters  $u=(u_1,u_2, \dots ,u_m)$, each of which can only take on a fixed (discrete) set of values:
\begin{equation}\label{eq:3}
U=U_1 \times U_2 \times \dots \times U_m=\{w_k=<w_{1k},w_{2k}, \dots, w_{mk}>: 1 \leq k \leq l\}, w_{ik} \in U_i, 
\end{equation}
where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_i$, i.e., the whole set of all possible discrete parameter values $U$ contains
\begin{equation}\label{eq:4}
  l=\prod\limits^m_{i = 1}{l_i}
\end{equation}
different elements (tuples $w_k$,$1 \leq k \leq l$). Without loss of generality, it will be assumed in further consideration that the criteria $f_i (y,u)$, $1 \leq i \leq s$, are non-negative and their reduction corresponds to an increase in the efficiency of the decisions selected.
\end{itemize}

In the most complex variant, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multiextremal, and the procedures for calculating their values can be computationally time-consuming.  It is also assumed that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition   
\begin{equation}\label{eq:5}
  |f_i (y_1,u_1 )-f_i (y_2, u_2 )| \leq L_i \|(y_1, u_1)-(y_2, u_2)\|, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the criterion $f_i (y,u)$, $1 \leq i \leq s$,  and $\|*\|$ denotes the Euclidean norm in $R^N$.

Efficiency criteria in an MCO problem are usually controversial and the parameters $(y^*,u^*) \in D \times U$ with the optimal values for all criteria at the same time may not be available. In such situations, it is a common approach for MCO problems to find efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of efficiency indicators for other criteria. Obtaining the whole set of efficient decisions (the Pareto set) can require a lot of computation and, as a result, another approach is often used -- to find only a relatively small set of efficient decisions defined according to the decision maker's requirements.

A commonly used approach to obtaining efficient decisions is to transform a vector criterion into some combined scalar efficiency function\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
\begin{equation}\label{eq:6}
  \min{F(\alpha,y,u)}, y \in D, u \in U,
\end{equation}
where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of the parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2})-(\ref{eq:3}). By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, i.e.  
\begin{equation}\label{eq:7}
  |F(\alpha, y_1, u_1)-F(\alpha, y_2, u_2)| \leq L\|(y_1,u_1)-(y_2,u_2)\|.
\end{equation}

To construct a combined scalar efficiency function $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
\begin{equation}\label{eq:8}
\begin{matrix}
  F(\lambda,y,u)=\max{(\lambda_i f_i (y,u),1\leq i \leq s)},\\
  \lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{matrix}
\end{equation}

It should be noted that due to the possible changes in the requirements for optimality in the process of calculations it may be necessary to change the parameters of the convolution $\lambda$ from (\ref{eq:8}). Such variations yield a set of scalar global optimization problems (\ref{eq:6})
\begin{equation}\label{eq:10}
\mathbb{F}_T=\{F(\alpha_t,y , u):1 \leq t \leq T\},
\end{equation}
which is necessary for solving the MCOmix problem. This set of problems can be formed sequentially in the course of calculations; the problems of the set can be solved strictly sequentially or simultaneously in time-shared mode. In addition, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems. The possibility of forming the set $\mathbb{F}_T$ defines a \textit{new approach to multistage solution of multicriteria optimization} (MMCO) problems -- see, for example, \cite{c22}.

@@

In the presence of discrete parameters, solving multicriteria optimization  problems becomes considerably  more complicated -- in many cases, it becomes necessary to calculate criterion values for all possible values of discrete parameters. The proposed approach to improving the efficiency of solving MCOmix problems is based on the following two basic ideas: unrolling mixed-integer optimization problems \cite{c23} and dimensionality reduction \cite{c24,c25}.

@@

To solve the global optimization problem (\ref{eq:6}), a two-stage nested optimization scheme can be used
\begin{equation}\label{eq:11}
\begin{matrix}
  F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} = \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} = \\ 
  \min_{y \in D}(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )),\\
	w_i\in U,1 \leq i \leq l.
\end{matrix}
\end{equation}

In the computational scheme (\ref{eq:11}), for any value of continuous parameters $y \in D$, the values of the function $F(\alpha,y,u)$ are calculated for all possible values of discrete parameters $u \in U$. However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of discrete parameters $u^*$, so the calculation of function $F(\alpha,y,u)$ values for other values of discrete parameters is redundant. Thus, for example, for the problem
\begin{equation}\label{eq:12}
 \min\{u^2(sin(x)+sin(10x/3)):x\in[2.7,7.5],u\in \{1,2\}\}
\end{equation}
having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form shown in Fig.~\ref{fig:1} and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig1}
  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=2$)}
  \label{fig:1}
\end{figure}

It is possible to increase the efficiency and, accordingly, to reduce computational complexity when solving  the problem (\ref{eq:6}) by excluding (or at least by reducing) the calculation of the function $F(\alpha,y,u)$ for discrete parameters $u \neq u^*$. In some rare situations, the information about the values of the discrete parameters that do not allow the minimum value of the function $F(\alpha,y,u)$ to be reached, may be known a priori. In most cases, such information can be obtained only in the process of calculations based on the calculated search information.

Assume that 
\begin{equation}\label{eq:13}
 \Omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ): 1\leq i\leq k \}
\end{equation}
is the search information obtained in the process of calculation after $k>1$ global search iterations. Then the procedure for adaptive estimation of the discrete parameter value, at which the minimum value  of the function $F(\alpha,y,u)$ is expected to be obtained, can be determined using the decision rule
\begin{equation}\label{eq:14}
	\theta =\Theta(\Omega_k ),1 \leq \theta \leq l, 
\end{equation}
 that determines the most promising value of the discrete parameter $u \in U$ at each step of the global search.

@@

\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\Omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.

\textit{Rule 2.} Determine the point $y \in D$ of the global search iteration to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.

\textit{Rule 3.} Check the stopping condition for the computation. If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\Omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations starting from Rule 1.

@@

It should be noted that the computational complexity of solving global optimization problems increases exponentially with increasing dimensionality. The problem of increasing computational complexity was even called the ``dimensionality curse''. In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\Omega_k$ from (\ref{eq:13}). This computational complexity can be significantly reduced by reducing the optimization problems to be solved by using Peano \textit{curves} or \textit{evolvents} $y(x)$, which unmbiguously and continuously map a segment [0,1] on an $n$-dimensional hyper-cube $D$ -- see, for example, \cite{c24,c25}. As a result of this reduction, the multidimensional global optimization problem (\ref{eq:6}) is reduced to a one-dimensional problem
\begin{equation}\label{eq:15}
 \min_{y\in D, u \in U}F(\alpha,y,u) = \min_{x \in [0,1]}(F(\alpha,y(x),w_1 ),F(\alpha,y(x),w_2 ),\dots,F(\alpha,y(x),w_l ))
\end{equation}

It should be noted that the one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction satisfy the uniform H{\" o}lder condition, i.e.
\begin{equation}\label{eq:16}
|F(\alpha, y(x_1),u_1)-F(\alpha, y(x_2),u_2)| \leq H|x_1-x_2 |^{1/n},
\end{equation}
where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}) and $n$ is the dimensionality of the MCOmix problem from (\ref{eq:1}).

As a result of dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:13}) obtained in the calculation can be presented in the form 
\begin{equation}\label{eq:17}
A_k=\{(x_i, u_i, z_i, f_i = f(y(x_i), u_i): 1 \leq i \leq k \},
\end{equation}
where $x_i$, $u_i$, $1 \leq i \leq k$,  are the points of the global search iterations, $z_i$, $f_i$, $1 \leq i \leq k$,  are the values of the scalar criterion $F(\alpha,y(x),u)$ and the criteria $f_i (y)$, $1 \leq i \leq s$, calculated at the points $x_i$, $u_i$, $1 \leq i \leq k$. Note that the data in the set $A_k$ are arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of the points $x_i$, $1 \leq i \leq k$, i.e.
\begin{equation}\label{eq:18}
 x_1< x_2< \dots < x_k
\end{equation}
for a more efficient execution of global search algorithms.

The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions $F(\alpha, y(x),w_i)$, $1 \leq i \leq l$, from (\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).

\begin{equation}\label{eq:19}
 \Phi(\alpha,y(x)) = 
\begin{cases}
F(\alpha,y(x),w_1 ),x \in [0,1], \\
F(\alpha,y(x),w_2 ),x \in (1,2], \\
\dots \\
F(\alpha,y(x),w_l ),x \in (l-1,l],
\end{cases}
\end{equation}
where $l$ from (\ref{eq:4}) is the number of different variants of the values of discrete parameters $u \in U$.  The mapping of the extended segment $[0,l]$ onto the domain $D$ of the continuous parameter values from (\ref{eq:2}) can be determined as follows 
\begin{equation}\label{eq:20}
Y(x)=y(x-E(x)),x\in[0,l]
\end{equation}
($E(x)$ denotes the operation of taking the integral part of the number $x$). It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined and are not used in the calculations.

@@

In the general case, the problem of minimizing the function $F(\lambda,y,u)$ from (\ref{eq:8}) is one of  global optimization problems. The solution of such problems involves the construction of grids covering the search domain $D$ -- see, for example, \cite{c24,c25,c26,c27,c28,c29,c30,c31}.

The proposed approach uses the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$. This algorithm expands the possibilities of multiextremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).

The AGMIS general computational scheme can be presented as follows (see also \cite{c21}).

>>>>

\makeatletter\@ifundefined{UseRawInputEncoding}{\newcommand\UseRawInputEncoding{}}{\UseRawInputEncoding}\makeatother
\documentclass{svproc}

@@

\begin{abstract}
The paper discusses a new approach to solving computationally time-consuming multicriteria optimization problems in which some variable parameters can only take on discrete values. Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of optimization problems where only continuous parameters are used. All problems of the family are solved simultaneously in time-shared mode, where the optimization problem for the next global search iteration is selected adaptively, taking into account the search information obtained in the course of the calculations. The suggested algorithms enable parallel computing on high-performance computing systems. The computational experiments confirm that the proposed approach can significantly reduce the computation volume and time required for solving complex multicriteria mixed-integer optimization problems.

\keywords{Multicriteria optimization $\cdot$ Mixed-integer optimization problems $\cdot$ Methods of criteria scalarization $\cdot$ Global optimization $\cdot$ Search information $\cdot$ Parallel computing $\cdot$ Computational experiment}
\end{abstract}

@@

Multicriteria optimization (MCO) problems arise whenever optimal decisions need to be made during the development of complex technical devices and systems. The scale of demand for MCO problems determines the high intensity of research in this area (see, for example, the monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}).

Usually, the solution to an MCO problem is a set of efficient (non-dominated) decisions which cannot be improved under some criteria without deteriorating the efficiency under some other criteria. The definition of the whole set of efficient decisions (the Pareto set) may, on the one hand, require a large amount of computation and, on the other, may be redundant as the analysis of a large number of efficient decisions may require a significant effort from the person who makes the decision (the decision maker, DM). Thus, it may be practically justified to find only a relatively small set of efficient decisions that can be formed according to the optimality requirements defined by the DM.

By restricting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the number of computations required. However, efficiency criteria may be complex, \textit{multiextremal}, and calculating the values of these criteria can prove \textit{computationally demanding}. Besides, some variables may only take on discrete values. In such cases, MCO problems involve a \textbf{significant computational complexity}, which can only be overcome by using high-performance supercomputer systems.

Many different approaches have been proposed for solving MCO problems (see, for example, \cite{c3,c6,c10,c11}.  Most commonly, various methods based on the reduction of the vector criterion to a particular scalar function are used \cite{c2,c12}. The number of works concerned with multicriteria mixed-integer problems is however more limited; in most cases, the issues of discrete parameter analysis are considered in connection with scalar optimization problems (see, for example, reviews \cite{c13,c14}). The widely used deterministic methods for solving problems of this class are usually based on either the Branch-and-Bound \cite{c15} or the Branch-and-Reduce approaches \cite{c16}. Several meta-heuristic and genetic algorithms are also known and are based, in one way or another, on the random search concept \cite{c17,c18}.

This paper presents the results of a study concerned with the development of highly effective parallel methods of multicriteria optimization making use of all the search information obtained throughout computations \cite{c19,c20,c21}. A new contribution to this research field is the development of an approach for solving MCO problems in which some of the varied parameters can only take on discrete values. The suggested approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used. All problems of the family are solved simultaneously in time-shared mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively, taking into account the search information obtained in the course of computations. The developed algorithms enable efficient parallel computing on high-performance computing systems.

The rest of the paper is structured as follows. In Section \ref{sec:2}, we statement of multicriteria optimization problems, present a minimax scheme for scalarization of the vector efficiency criterion, and introduce the concept of multistage solution for multicriteria optimization problems. In Section \ref{sec:3}, we describe the proposed approach based on the reduction of mixed-integer optimization problems to the solution of a family of optimization problems using only continuous parameters. In that section, we also describe a dimensionality reduction scheme by which multidimensional optimization problems can be reduced to one-dimensional global search problems. Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems under of the proposed approach. Section \ref{sec:5} contains the results of numerical experiments that confirm that the approach we propose is promising. In the Conclusions section, we discuss the results obtained and outline possible directions for future research.

@@

The problem of multicriteria mixed-integer optimization (MCOmix) can be expressed as follows
\begin{equation}\label{eq:1}
f(y,u)\to \min,\ y\in D,\ u \in U,
\end{equation}
where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency, in which the varied parameters belong to two different types:
\begin{itemize}
	\item continuous parameters $y=(y_1,y_2,\dots,y_n)$, whose domain of possible values is represented by an $N$-dimensional hyperparallelepiped,
\begin{equation}\label{eq:2}
D=\{ y\in R^n: a_i \leq y_i \leq b_i, 1 \leq i \leq n \},
\end{equation}
for specified vectors $a$ and $b$;

	\item discrete parameters  $u=(u_1,u_2, \dots ,u_m)$, each of which can only take on a fixed (discrete) set of values,
\begin{equation}\label{eq:3}
U=U_1 \times U_2 \times \dots \times U_m=\{w_k=\langle w_{1k},w_{2k}, \dots, w_{mk}\rangle : 1 \leq k \leq l\},\ w_{ik} \in U_i,
\end{equation}
where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_i$, i.e., the set $U$ of all possible discrete parameter values contains
\begin{equation}\label{eq:4}
  l=\prod\limits^m_{i = 1}{l_i}
\end{equation}
different elements (tuples $w_k$, $1 \leq k \leq l$). Without loss of generality, we assume below that the criteria $f_i(y,u)$, $1 \leq i \leq s$, are non-negative, and their reduction corresponds to an increase in the efficiency of the decisions selected.
\end{itemize}

In the most difficult case, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multiextremal, and the procedures for calculating their values can be computationally time-consuming.  We also assume that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition
\begin{equation}\label{eq:5}
  |f_i (y_1,u_1 )-f_i (y_2, u_2 )| \leq L_i \|(y_1, u_1)-(y_2, u_2)\|, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the criterion $f_i (y,u)$, $1 \leq i \leq s$,  and $\|*\|$ denotes the Euclidean norm in $R^N$.

Efficiency criteria in an MCO problem are usually controversial, and there may not exist parameters $(y^*,u^*) \in D \times U$ with values simultaneously optimal for all criteria. In such situations, it is a common practice in MCO problems to search for efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of the efficiency indicators for other criteria. Obtaining the whole set of efficient decisions (the Pareto set) can require a lot of computation and, for this reason, a different approach is often used, namely search only for a relatively small set of efficient decisions defined according to the decision maker's requirements.

A commonly used approach to obtaining efficient decisions is to transform the vector criterion into some combined scalar efficiency function\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
\begin{equation}\label{eq:6}
  \min{F(\alpha,y,u)},\ y \in D,\ u \in U,
\end{equation}
where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2})--(\ref{eq:3}). By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, that is,
\begin{equation}\label{eq:7}
  |F(\alpha, y_1, u_1)-F(\alpha, y_2, u_2)| \leq L\|(y_1,u_1)-(y_2,u_2)\|.
\end{equation}

To construct a combined scalar efficiency function $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
\begin{equation}\label{eq:8}
\begin{matrix}
  F(\lambda,y,u)=\max{(\lambda_i f_i (y,u),1\leq i \leq s)},\\
  \lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0,\ 1 \leq i \leq s.
\end{matrix}
\end{equation}

It should be noted that, due to the possible changes in the requirements for optimality in the process of calculation, it may be necessary to change the parameters of the convolution $\lambda$ from (\ref{eq:8}). Such variations yield a set of scalar global optimization problems (\ref{eq:6})
\begin{equation}\label{eq:10}
\mathbb{F}_T=\{F(\alpha_t,y , u):1 \leq t \leq T\},
\end{equation}
which is necessary for solving the MCOmix problem. This set of problems can be formed sequentially during the calculations; the problems of the set can be solved strictly sequentially or simultaneously in time-shared mode. Furthermore, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems. The possibility of forming the set $\mathbb{F}_T$ determines \textit{a new approach to the multistage solution of multicriteria optimization} (MMCO) problems (see, for instance, \cite{c22}).

@@

The solution of multicriteria optimization  problems becomes considerably  more complicated in the presence of discrete parameters: in many cases, it is necessary to calculate the criterion values for all possible values of the discrete parameters. The proposed approach to improving the efficiency of the solution of MCOmix problems is based on two basic ideas: unrolling mixed-integer optimization problems \cite{c23} and dimensionality reduction \cite{c24,c25}.

@@

To solve the global optimization problem (\ref{eq:6}), a two-stage nested optimization scheme can be used:
\begin{equation}\label{eq:11}
\begin{gathered}
  F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} = \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} = \\
  \qquad=\min_{y \in D}\bigl(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )\bigr),\\
	w_i\in U,\ 1 \leq i \leq l.
\end{gathered}
\end{equation}

In computational scheme (\ref{eq:11}), the values of the function $F(\alpha,y,u)$ for any value of continuous parameters $y \in D$ are computed for all possible values of the discrete parameters $u \in U$. However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of the discrete parameters $u^*$, so the computation of the values of the function $F(\alpha,y,u)$ for other values of the discrete parameters is redundant. For example, consider the problem
\begin{equation}\label{eq:12}
 \min\{u^2(\sin(x)+\sin(10x/3)):x\in[2.7,7.5],\ u\in \{1,2\}\},
\end{equation}
which has one continuous and one discrete parameter. The plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have the form shown in Figure~\ref{fig:1} and, as we can see, there is no need to compute the value of the function $F(\alpha,y,u)$ for $u=1$.

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig1}
  \caption{The function $F(\alpha,y,u)$ from problem (\ref{eq:12}) plotted for different values of the discrete parameter (on the left, $u=1$;  on the right, $u=2$)}
  \label{fig:1}
\end{figure}

It is possible to increase the efficiency and, accordingly, to reduce the computational complexity when solving the problem (\ref{eq:6}), by excluding (or, at least, by reducing) the computation of the function $F(\alpha,y,u)$ for discrete parameters $u \neq u^*$. In some rare situations,  we may know a priori information about the values of discrete parameters that do not allow the function $F(\alpha,y,u)$ to reach a minimum value. In most cases, that information can be obtained only in the process of computations based on the calculated search information.

Assume that
\begin{equation}\label{eq:13}
 \Omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ) : 1\leq i\leq k \}
\end{equation}
is the search information obtained in the process of calculation after $k>1$ global search iterations. Then, the procedure for adaptive estimation of the discrete parameter value at which the function $F(\alpha,y,u)$ is expected to reach the minimum value can be determined using the decision rule
\begin{equation}\label{eq:14}
	\theta =\Theta(\Omega_k ),\ 1 \leq \theta \leq l,
\end{equation}
which determines the most promising value of the discrete parameter $u \in U$ at each step of the global search.

@@

\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to the available search information $\Omega_k$ from (\ref{eq:13})  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.

\textit{Rule 2.} Determine the point $y \in D$ of the global search iteration that needs to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.

\textit{Rule 3.} Check the stopping condition of the computation. If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\Omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue the computation starting from Rule~1.

@@

Note that the computational complexity of the solution of global optimization problems increases exponentially as the dimensionality grows. The problem of growth of the computational complexity was even called the ``dimensionality curse''. In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze the multidimensional search information $\Omega_k$ from (\ref{eq:13}). This computational complexity can be significantly reduced by reducing the optimization problems to be solved, through the use of \textit{Peano curves} or \textit{evolvents} $y(x)$ that unambiguously and continuously map the segment $[0,1]$ onto an $n$-dimensional hypercube $D$ (see, for example, \cite{c24,c25}). As a result, the multidimensional global optimization problem in (\ref{eq:6}) is reduced to a one-dimensional problem, namely
\begin{equation}\label{eq:15}
\min_{y\in D, u \in U}F(\alpha,y,u) = \min_{x \in [0,1]} \bigl(F(\alpha,y(x),w_1 ),F(\alpha,y(x),w_2 ),\dots,F(\alpha,y(x),w_l )\bigr)
\end{equation}

We should also note that one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction satisfy a uniform H{\"o}lder condition,
\begin{equation}\label{eq:16}
|F(\alpha, y(x_1),u_1)-F(\alpha, y(x_2),u_2)| \leq H \lvert x_1-x_2\rvert^{1/n},
\end{equation}
where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}), and $n$ is the dimensionality of the MCOmix problem from (\ref{eq:1}).

As a result of the dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:13}) obtained by computation can be expressed as
\begin{equation}\label{eq:17}
A_k=\{(x_i, u_i, z_i, f_i = f(y(x_i), u_i) : 1 \leq i \leq k \},
\end{equation}
where $x_i, u_i$, $1 \leq i \leq k$,  are the points of the global search iterations, and $z_i$, $f_i$, $1 \leq i \leq k$,  are the values of the scalar criterion $F(\alpha,y(x),u)$ and criteria the $f_i (y)$, $1 \leq i \leq s$, computed at the points $x_i$, $u_i$, $1 \leq i \leq k$. Note that the data in the set $A_k$ are arranged in ascending order\footnote{The order is indicated by the subscript.} of the points $x_i$, $1 \leq i \leq k$, that is,
\begin{equation}\label{eq:18}
 x_1< x_2< \dots < x_k,
\end{equation}
for more efficient execution of global search algorithms.

Dimensionality reduction makes it possible to combine (concatenate) the one-dimensional functions $F(\alpha, y(x),w_i)$, $1 \leq i \leq l$, from~(\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$ defined over the segment $[0,l]$ (see Figure~\ref{fig:1}),
\begin{equation}\label{eq:19}
 \Phi(\alpha,y(x)) =
\begin{cases}
F(\alpha,y(x),w_1 ),\ x \in [0,1], \\
F(\alpha,y(x),w_2 ),\ x \in (1,2], \\
\hdotsfor{1} \\
F(\alpha,y(x),w_l ),\ x \in (l-1,l],
\end{cases}
\end{equation}
where $l$, as expressed in (\ref{eq:4}), is the number of different variants of the values of discrete parameters $u \in U$.  The mapping of the extended segment $[0,l]$ onto the domain $D$ of continuous parameter values from (\ref{eq:2}) can be determined as
\begin{equation}\label{eq:20}
Y(x)=y(x-E(x)),\ x\in[0,l]
\end{equation}
where $E(x)$ denotes the integral part of the number $x$. Note that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$. Below, we consider the values of the function $\Phi(\alpha,y(x))$ at these points undefined and will not use them in computations.

@@

In the general case, the problem of minimizing the function $F(\lambda,y,u)$ from (\ref{eq:8}) is one of the global optimization problems. The solution to such problems involves the construction of grids covering the search domain $D$ (see, for example, \cite{c24,c25,c26,c27,c28,c29,c30,c31}).

The proposed approach relies on the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$. This algorithm expands the possibilities of multiextremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).

The AGMIS general computational scheme can be described as follows (see also \cite{c21}).
